---
title: "Project 2 - Predictive Models"
author: "Alex Prevatte & Chennade Brown"  
date: "10/31/2021"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Our project involves creating predictive models and automating Markdown reports for each article topic (Lifestyle, Social Media, Business, World, Entertainment, and Technology) using the [Online News Popularity Data Set](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity). This dataset contains quantitative information including the number of words, images, and links, as well as qualitative information such as article topic and day published. The goal of this report is to predict the number of shares for each article topic using linear regression, random forest, and boosted tree models. Our team used 55 prediction attributes for ensemble methods and a combination of the following attributes for linear regression to predict article shares:

`n_tokens_title` : Number of words in the title
`n_tokens_content` : Number of words in the content
`rate_positive_words` : Rate of positive words among non-neutral tokens
`num_hrefs` : Number of links
`num_imgs` : Number of images
`num_keywords` : Number of keywords in the metadata
`global_subjectivity` : Text subjectivity
`max_positive_polarity` : Max. polarity of positive words
`num_imgs` : Number of images
`kw_avg_avg` : Avg. keyword (avg. shares)
`kw_min_avg` : Avg. keyword (min. shares)
`kw_max_avg` : Avg. keyword (max. shares)

```{r, echo = TRUE, message = FALSE}
library(tidyverse)
library(dplyr)
library(knitr)
library(GGally)
library(corrplot)
library(caret)
library(randomForest)
```

## Data
```{r, echo = TRUE, message = FALSE}
# Import the Data.
 onlineNews <- read_csv("OnlineNewsPopularity.csv")
```

```{r, echo = TRUE}
# Start with using data from a single data_channel_is_* source

# Filtered rows for data_channel_is_lifestyle and removed other 5 categories. 
# Removed non-predictive variables (URL, Timedelta).

onlineNewsLifestyle <- onlineNews %>% filter(`data_channel_is_lifestyle` ==  1) %>% 
  select(- c(`data_channel_is_bus`, `data_channel_is_entertainment`,
             `data_channel_is_socmed`, `data_channel_is_tech`,
             `data_channel_is_world`, url, `timedelta`))

# Remove space in front of each column name
names(onlineNewsLifestyle) <- sub(" ", "", names(onlineNewsLifestyle))

# Convert to dataframe for plotting
onlineNewsLifestyle <- data.frame(onlineNewsLifestyle)
```

## Summarizations
```{r, echo = TRUE}
# Summary statistics for shares
sum <- summary(onlineNewsLifestyle$shares)
std <- sd(onlineNewsLifestyle$shares)
# Median and mean to be used for histogram plot
median <- sum[3]
mean <- sum[4]
print(sum)

# Summary statistics for Number of Words in Title
sum <- summary(onlineNewsLifestyle$n_tokens_title)
std <- sd(onlineNewsLifestyle$n_tokens_title)
print(sum)

# Summary statistics for Number of Words in Content
sum <- summary(onlineNewsLifestyle$n_tokens_content)
std <- sd(onlineNewsLifestyle$n_tokens_content)
print(sum)
```

```{r}
# Create Day variable to combine days for contingency table
dayShares <- onlineNewsLifestyle %>% 
  mutate(Day = if_else(weekday_is_monday == 1, "M",
                          if_else(weekday_is_tuesday == 1, "T",
                                  if_else(weekday_is_wednesday == 1, "W", 
                                          if_else(weekday_is_thursday == 1, "R",
                                                  if_else(weekday_is_friday == 1,"F",
                                                          if_else(weekday_is_saturday == 1,"Sat",
                                                                  if_else(weekday_is_sunday == 1,"Sun", "NULL"))))))))

# Add Popular variable for rows where shares is greater than the median
dayShares <- dayShares %>% mutate(Popular = 
                                    ifelse(shares > median, "Yes", "No")) %>% select(Popular, everything())

# Contingency table for day of the week and shares greater than the median
table(dayShares$Popular, factor(dayShares$Day, levels = c("M", "T", "W", 
                                                          "R", "F", "Sat", "Sun")))

# Table of shares for each day
tableTotal <- aggregate(dayShares$shares, by=list(Day=dayShares$Day), FUN=sum)
colnames(tableTotal) <- c("Day", "Total_Shares")
tableTotal$Day <- factor(tableTotal$Day, levels = c("M", "T", "W", 
                                                          "R", "F", "Sat", "Sun"))
tableTotal
      
```

```{r, echo = TRUE, warning = FALSE}
# Scatterplot of Number of Words in Title vs. Number of Shares
titleCorrelation <- cor(onlineNewsLifestyle$shares, onlineNewsLifestyle$n_tokens_title)

ggplot(onlineNewsLifestyle, aes(n_tokens_title, shares)) + 
  labs(y = "Number of Shares", x = "Number of Words in Title") +
  ggtitle("Number of Words in Title vs. Number of Shares") +
  geom_point() + 
  geom_text(x = 15, y = 200000, size = 5, 
            label = paste0("Correlation =", round(titleCorrelation, 3)))

# Histogram of Total Shares. Mean and Median vertical lines have been included.
ggplot(onlineNewsLifestyle, aes(x=shares)) +
  labs(y = "Frequency", x = "Number of Shares") +
  ggtitle("Frequency of Total Shares") +
  geom_histogram(binwidth = 500, color="darkblue", fill="lightblue") +
  xlim(0,30000) +
  geom_vline(aes(xintercept = median), colour="black") + 
  geom_vline(aes(xintercept = mean), colour="red")

# Barplot of Total Shares vs. Day of the Week 
ggplot(tableTotal, aes(x=Day, y = Total_Shares)) +
  geom_bar(stat="identity", fill = "steelblue") +
  labs(y = "Number of Shares", x = "Day of the Week") +
  ggtitle("Total Shares by Day of the Week") 
```

```{r, echo = TRUE}
# Summary statistics for the number of words in the content.
sumWords <- summary(onlineNewsLifestyle$n_tokens_content)
sumWords
```

```{r, echo = TRUE}
# Summary statistics for the number of links in the content. 
sumLinks <- summary(onlineNewsLifestyle$num_hrefs)
sumLinks
```

```{r, echo = TRUE}
# Summary statistics for the number of images in the content.
sumImages <- summary(onlineNewsLifestyle$num_imgs)
sumImages
```

```{r, echo = TRUE}
# Add Word variable for words to include low, average, and high word count and combine with dayshares dataframe for contingency table.
wordLinkShares <- dayShares %>% mutate(Words = if_else(n_tokens_content <= 500, "Low", if_else(n_tokens_content <= 625, "Average", "High")))

```

```{r, echo = TRUE}
# Add Link variable for the number of links in the content to include low, average, and high count.
wordLinkShares <- wordLinkShares %>% mutate(Links = if_else(num_hrefs <= 9, "Low", if_else(num_hrefs <= 14, "Average","High")))

# Remove extra columns added to onLineNewsLifestyle.
onlineNewsLifestyle$WordCount <- NULL
onlineNewsLifestyle$Popular <- NULL
onlineNewsLifestyle$Words <- NULL
onlineNewsLifestyle$Links <- NULL
onlineNewsLifestyle$Day <- NULL
```

```{r, echo = TRUE}
# Contingency table for the number of words in the content based on grouping the word count into categories of low, average, and high word count and grouping the shares based on popularity (shares greater than the median).
table(wordLinkShares$Popular, factor(wordLinkShares$Words, levels = c("Low", "Average", "High")))
```

```{r, echo = TRUE}
# Contingency table for the number of links in the content based on grouping the link count into categories of low, average,and high and shares based on popularity (shares greater than the median)
table(wordLinkShares$Popular, factor(wordLinkShares$Links, levels = c("Low", "Average", "High")))
```

```{r, echo = TRUE, message = FALSE}
# Select predictors to view in GGPairs plot.
xpred <- onlineNewsLifestyle %>% select(kw_avg_min, n_tokens_content, num_videos, avg_negative_polarity, rate_negative_words, shares)

xpred2 <- onlineNewsLifestyle %>% select(num_keywords, global_subjectivity, rate_positive_words, n_tokens_title, max_positive_polarity, shares)

# GGPairs plot to view correlation among the predictors.  Correlation greater than 75% indicates the predictors are highly correlated.
ggpairs(xpred, title = "Correlogram with ggpairs")
ggpairs(xpred2, title = "Correlogram with ggpairs")
```

```{r, echo = TRUE}
# The following scatterplot shows the trend of shares as a function of the number of links in the content.  An upward trend in the points indicates that articles with more links are shared more often.  A downward trend would indicate that articles with more links are shared less often.  If there is neither an upward or downward trend this indicates that the number of links in the article has no effect on whether the article will be shared.
correlation <- cor(onlineNewsLifestyle$shares, onlineNewsLifestyle$num_hrefs)  
g <- ggplot(onlineNewsLifestyle, aes(x = num_hrefs, y = shares)) + labs(y ="Number of Shares", x = "Number of Links")
g + geom_point(col = "purple") + ggtitle("Number of Links vs. Shares") + geom_text(x = 125, y = 30000, size = 5, label = paste0("Correlation = ", round(correlation, 2)))

```

```{r, echo = TRUE}
# The following scatterplot shows the relationship between the rate of positive words in the articles and the number of shares.  If the plots are on an upward trajectory then articles with more positive words are shared the most.  If the plots are on a downward trend then the articles with the most positive words are shared the least.
correlationTwo <- cor(onlineNewsLifestyle$shares, onlineNewsLifestyle$rate_positive_words)  
g <- ggplot(onlineNewsLifestyle, aes(x = rate_positive_words, y = shares)) + labs(y ="Number of Shares", x = "Rate of Positive Words")
g + geom_point(col = "blue") + ggtitle("Rate of Positive Words vs. Shares") + geom_text(x = 0.75, y = 100000, size = 5, label = paste0("Correlation = ", round(correlationTwo, 2)))

```

## Modeling
```{r, echo = TRUE}
# Split the data into a training (70% of the data) and test set (30% of the data).
set.seed(90)
trainIndex <- createDataPartition(onlineNewsLifestyle$shares, p = 0.7, list = FALSE)
onlineNewsTrain <- onlineNewsLifestyle[trainIndex, ]
onlineNewsTest <- onlineNewsLifestyle[-trainIndex, ]
dim(onlineNewsTrain)
dim(onlineNewsTest)
```

### Linear Regression Models
A linear regression model is a model used to determine the relationship between two or more variables by fitting a linear equation to the data.  The linear equation is in the form of: Y = Bo + B1X + E where Y is the response variable, X is the explanatory variable, Bo is the y intercept, and B1 is the slope.  The model can be used to predict the value of the response variable based on the values of the explanatory variables.  For example a linear regression model can be used to model the relationship between years of experience and salary.
```{r, echo = TRUE}
# Store model one in a formula.
modelOne <- as.formula("shares ~ n_tokens_content + n_tokens_title + rate_positive_words + num_keywords + global_subjectivity + max_positive_polarity + num_imgs")

# Model fit on the training data.
fit1 <- train(modelOne, data = onlineNewsTrain,
              method = "lm",
              preProcess = c("center", "scale"),
              trControl = trainControl(method = "cv", number = 10))
fit1

# Backward selection chosen for additional linear model
modelTwo <- as.formula("shares ~ kw_avg_avg + kw_min_avg + kw_max_avg + num_hrefs + n_tokens_content")

# Model fit on training data
fit2 <- train(modelTwo, data = onlineNewsTrain,
              method = "lm",
              preProcess = c("center", "scale"),
              trControl = trainControl(method = "cv", number = 10))
fit2
```

### Random Forest Model
Random Forest models are an extension of the tree based method bagging.  The random forest algorithm creates multiple trees from bootstrapped samples, averages those results, and uses a random subset of predictors for each bootstrap sample/tree fit.  Random forests can be used for classification and regression problems.

```{r, echo = TRUE}
rfFit <- train(shares ~., data = onlineNewsTrain, 
               method = "rf", 
               trainControl = trainControl(method = "cv",
                                           number = 5),
               tuneGrid = data.frame(mtry = 1:10))
rfFit
```

### Boosted Tree Model
Boosting is an ensemble tree-based method used for regression and classification. Boosting seeks to improve prediction by training on decision trees which are grown sequentially. Each subsequent tree is grown on a modified version of the original data. The predictions are then updated as new trees are grown, minimizing the error of the previous tree. The parameters used in boosting are the number of trees, interaction depth (number of splits on a tree), shrinkage, and minimum number of observations in trees' terminal nodes.  

```{r, results = 'hide'}
# Boosted tree. After examining ranges of values for number of trees and interaction depth, 25 trees with an interaction depth of 1 had the lowest test RMSE.
boostedGrid <- expand.grid(
  interaction.depth=c(1:4), 
  n.trees=c(25,50,100,150,200),
  shrinkage=0.1,
  n.minobsinnode=10
)

boostedTree <- train(shares ~. -data_channel_is_lifestyle, 
                     data = onlineNewsTrain,
             method = "gbm",
             preProcess = c("center", "scale"),
             trControl = trainControl(method = "repeatedcv", number = 5,
                                      repeats = 3),
             tuneGrid = boostedGrid)
```

## Comparison
```{r, echo = TRUE}
# Test error computed for the first linear model.
pred <- predict(fit1, newdata = onlineNewsTest)
m <- postResample(pred, obs = onlineNewsTest$shares)

# Test error computed for the backwards selection linear model.
pred2 <- predict(fit2, newdata = onlineNewsTest)
p <- postResample(pred2, obs = onlineNewsTest$shares)

# Test error computed for the random forest model.
predForest <- predict(rfFit, newdata = onlineNewsTest)
n <- postResample(predForest, onlineNewsTest$shares)

# Test error computed for the boosted tree model.
predBoost <- predict(boostedTree, newdata = onlineNewsTest)
p2 <- postResample(predBoost, obs = onlineNewsTest$shares)


# Produce a table for all comparisons
comparisons <- t(rbind(p[1],p2[1],m[1],n[1]))
colnames(comparisons) <- c("Linear Model Backwards Selection", "Boosted Tree", "Linear Model One", "Random Forest")
comparisons
```

```{r, echo = TRUE}
# Compare RMSE values and store in a data frame..
LmOneRmse <- sqrt(mean((pred-onlineNewsTest$shares)^2))
LmBackSel <- sqrt(mean((pred2-onlineNewsTest$shares)^2))
rForestRmse <- sqrt(mean((predForest-onlineNewsTest$shares)^2))
tree <- sqrt(mean((predBoost-onlineNewsTest$shares)^2))
modelRmse <- c(LinearModelOne = LmOneRmse, LinearModelBckSel = LmBackSel, rForest = rForestRmse, boostedTree = tree)

# Store results in a data frame.
df <- as.data.frame(modelRmse)

# Name the column in the data frame.
attributes(df)$names[1] <- "RMSE"
df

# Use the slice_min function to return the row with the lowest RMSE.
BestModel <- df %>% slice_min(RMSE)
BestModel

# Use the apply function to return the final winner.
apply(X = select(df, RMSE), MARGIN = 2,
      FUN = function(x){
        winner <- df %>% slice_min(RMSE)
        names(winner) <- c("Winner")
        winner
      })

```



## Automation



