---
title: "Project 2 - Predictive Models"
author: "Alex Prevatte & Chennade Brown"  
date: "10/31/2021"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Our project involves creating predictive models and automating Markdown reports for each article topic (Lifestyle, Social Media, Business, World, Entertainment, and Technology) using the [Online News Popularity Data Set](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity). This dataset contains quantitative information for each article including the number of words, images, and links as well as qualitative information such as article topic and day published. The goal of this report is to predict article popularity by the number of shares for each topic using linear regression, random forest, and boosted tree models. Our team used the following attributes to predict article shares within each article topic:

* Note: Will update this based on models we decide to use

n_tokens_title
n_tokens_content
weekday_is_*
num_hrefs
num_imgs
.
.
.


```{r, echo = TRUE, message = FALSE}
library(tidyverse)
library(dplyr)
library(knitr)
library(GGally)
library(corrplot)
library(caret)
library(randomForest)
```

## Data
```{r, echo = TRUE, message = FALSE}
# Import the Data.
 onlineNews <- read_csv("OnlineNewsPopularity.csv")
```

```{r, echo = TRUE}
# Start with using data from a single data_channel_is_* source

# Filtered rows for data_channel_is_lifestyle and removed other 5 categories. 
# Removed non-predictive variables (URL, Timedelta).

onlineNewsLifestyle <- onlineNews %>% filter(`data_channel_is_lifestyle` ==  1) %>% 
  select(- c(`data_channel_is_bus`, `data_channel_is_entertainment`,
             `data_channel_is_socmed`, `data_channel_is_tech`,
             `data_channel_is_world`, url, `timedelta`))

# Remove space in front of each column name
names(onlineNewsLifestyle) <- sub(" ", "", names(onlineNewsLifestyle))

# Convert to dataframe for plotting
onlineNewsLifestyle <- data.frame(onlineNewsLifestyle)
```

## Summarizations

```{r, echo = TRUE}
# Summary statistics for shares
sum <- summary(onlineNewsLifestyle$shares)
std <- sd(onlineNewsLifestyle$shares)
# Median and mean to be used for histogram plot
median <- sum[3]
mean <- sum[4]
print(sum)

# Summary statistics for Number of Words in Title
sum <- summary(onlineNewsLifestyle$n_tokens_title)
std <- sd(onlineNewsLifestyle$n_tokens_title)
print(sum)

# Summary statistics for Number of Words in Content
sum <- summary(onlineNewsLifestyle$n_tokens_content)
std <- sd(onlineNewsLifestyle$n_tokens_content)
print(sum)
```

```{r}
# Create Day variable to combine days for contingency table
dayShares <- onlineNewsLifestyle %>% 
  mutate(Day = if_else(weekday_is_monday == 1, "M",
                          if_else(weekday_is_tuesday == 1, "T",
                                  if_else(weekday_is_wednesday == 1, "W", 
                                          if_else(weekday_is_thursday == 1, "R",
                                                  if_else(weekday_is_friday == 1,"F",
                                                          if_else(weekday_is_saturday == 1,"Sat",
                                                                  if_else(weekday_is_sunday == 1,"Sun", "NULL"))))))))

# Add Popular variable for rows where shares is greater than the median
dayShares <- dayShares %>% mutate(Popular = 
                                    ifelse(shares > median, "Yes", "No")) %>% select(Popular, everything())

# Contingency table for day of the week and shares greater than the median
table(dayShares$Popular, factor(dayShares$Day, levels = c("M", "T", "W", 
                                                          "R", "F", "Sat", "Sun")))

# Table of shares for each day
tableTotal <- aggregate(dayShares$shares, by=list(Day=dayShares$Day), FUN=sum)
colnames(tableTotal) <- c("Day", "Total_Shares")
tableTotal$Day <- factor(tableTotal$Day, levels = c("M", "T", "W", 
                                                          "R", "F", "Sat", "Sun"))
tableTotal
      
```

```{r, echo = TRUE}
# Scatterplot of Number of Words in Title vs. Number of Shares
titleCorrelation <- cor(onlineNewsLifestyle$shares, onlineNewsLifestyle$n_tokens_title)

ggplot(onlineNewsLifestyle, aes(n_tokens_title, shares)) + 
  labs(y = "Number of Shares", x = "Number of Words in Title") +
  ggtitle("Number of Words in Title vs. Number of Shares") +
  geom_point() + 
  geom_text(x = 15, y = 200000, size = 5, 
            label = paste0("Correlation =", round(titleCorrelation, 3)))

# Histogram of Total Shares. Mean and Median vertical lines have been included.
ggplot(onlineNewsLifestyle, aes(x=shares)) +
  labs(y = "Frequency", x = "Number of Shares") +
  ggtitle("Frequency of Total Shares") +
  geom_histogram(binwidth = 500, color="darkblue", fill="lightblue") +
  xlim(0,30000) +
  geom_vline(aes(xintercept = median), colour="black") + 
  geom_vline(aes(xintercept = mean), colour="red")

# Barplot of Total Shares vs. Day of the Week 
ggplot(tableTotal, aes(x=Day, y = Total_Shares)) +
  geom_bar(stat="identity", fill = "steelblue") +
  labs(y = "Number of Shares", x = "Day of the Week") +
  ggtitle("Total Shares by Day of the Week") 
```

```{r, echo = TRUE}
# Summary statistics for the number of words in the content.
sumWords <- summary(onlineNewsLifestyle$n_tokens_content)
sumWords
```

```{r, echo = TRUE}
# Summary statistics for the number of links in the content. 
sumLinks <- summary(onlineNewsLifestyle$num_hrefs)
sumLinks
```

```{r, echo = TRUE}
# Summary statistics for the number of images in the content.
sumImages <- summary(onlineNewsLifestyle$num_imgs)
sumImages
```

```{r, echo = TRUE}
# Add Word variable for words to include low, average, and high word count and combine with dayshares dataframe for contingency table.
wordLinkShares <- dayShares %>% mutate(Words = if_else(n_tokens_content <= 500, "Low", if_else(n_tokens_content <= 625, "Average", "High")))

```

```{r, echo = TRUE}
# Add Link variable for the number of links in the content to include low, average, and high count.
wordLinkShares <- wordLinkShares %>% mutate(Links = if_else(num_hrefs <= 9, "Low", if_else(num_hrefs <= 14, "Average","High")))

# Remove extra columns added to onLineNewsLifestyle.
onlineNewsLifestyle$WordCount <- NULL
onlineNewsLifestyle$Popular <- NULL
onlineNewsLifestyle$Words <- NULL
onlineNewsLifestyle$Links <- NULL
onlineNewsLifestyle$Day <- NULL
```

```{r, echo = TRUE}
# Contingency table for the number of words in the content based on grouping the word count into categories of low, average, and high word count and grouping the shares based on popularity (shares greater than the median).
table(wordLinkShares$Popular, wordLinkShares$Words)
```

```{r, echo = TRUE}
# Contingency table for the number of links in the content based on grouping the link count into categories of low, average,and high and shares based on popularity (shares greater than the median)
table(wordLinkShares$Popular, wordLinkShares$Links)
```

```{r, echo = TRUE, message = FALSE}
# Select predictors to view in GGPairs plot.
xpred <- onlineNewsLifestyle %>% select(n_tokens_content, n_tokens_title, num_keywords, rate_positive_words, rate_negative_words, shares)

xpred2 <- onlineNewsLifestyle %>% select(num_hrefs, global_subjectivity, num_imgs, title_subjectivity, max_positive_polarity, shares)

# GGPairs plot to view correlation among the predictors.  Correlation greater than 75% indicates the predictors are highly correlated.
ggpairs(xpred, title = "Correlogram with ggpairs")
ggpairs(xpred2, title = "Correlogram with ggpairs")
```

```{r, echo = TRUE}
# The following scatterplot shows the trend of shares as a function of the number of links in the content.  An upward trend in the points indicates that articles with more links are shared more often.  A downward trend would indicate that articles with more links are shared less often.  If there is neither an upward or downward trend this indicates that the number of links in the article has no effect on whether the article will be shared.
correlation <- cor(onlineNewsLifestyle$shares, onlineNewsLifestyle$num_hrefs)  
g <- ggplot(onlineNewsLifestyle, aes(x = num_hrefs, y = shares)) + labs(y ="Number of Shares", x = "Number of Links")
g + geom_point(col = "red") + ggtitle("Number of Links vs. Shares") + geom_text(x = 125, y = 30000, size = 5, label = paste0("Correlation = ", round(correlation, 2)))

```

```{r, echo = TRUE}
# The following scatterplot shows the relationship between the rate of positive words in the articles and the number of shares.  If the plots are on an upward trajectory then articles with more positive words are shared the most.  If the plots are on a downward trend then the articles with the most positive words are shared the least.
correlationTwo <- cor(onlineNewsLifestyle$shares, onlineNewsLifestyle$rate_positive_words)  
g <- ggplot(onlineNewsLifestyle, aes(x = rate_positive_words, y = shares)) + labs(y ="Number of Shares", x = "Rate of Positive Words")
g + geom_point(col = "red") + ggtitle("Rate of Positive Words vs. Shares") + geom_text(x = 0.75, y = 100000, size = 5, label = paste0("Correlation = ", round(correlationTwo, 2)))

```

## Modeling
```{r, echo = TRUE}
# Split the data into a training (70% of the data) and test set (30% of the data).
set.seed(90)
trainIndex <- createDataPartition(onlineNewsLifestyle$shares, p = 0.7, list = FALSE)
onlineNewsTrain <- onlineNewsLifestyle[trainIndex, ]
onlineNewsTest <- onlineNewsLifestyle[-trainIndex, ]
dim(onlineNewsTrain)
dim(onlineNewsTest)
```

### Linear Regression Models
A linear regression model is a model used to determine the relationship between two or more variables by fitting a linear equation to the data.  The linear equation is in the form of: Y = Bo + B1X + E where Y is the response variable, X is the explanatory variable, Bo is the y intercept, and B1 is the slope.  The model can be used to predict the value of the response variable based on the values of the explanatory variables.  For example a linear regression model can be used to model the relationship between years of experience and salary.

```{r, echo = TRUE}
# Store model one in a formula.
modelOne <- as.formula("shares ~ n_tokens_content + n_tokens_title + rate_positive_words + num_keywords + global_subjectivity + max_positive_polarity + num_imgs")

# Fit the linear regression models.
fit1 <- train(modelOne, data = onlineNewsTrain,
              method = "lm",
              preProcess = c("center", "scale"),
              trControl = trainControl(method = "cv", number = 10))
fit1
```

### Random Forest Model
Random Forest models are an extension of the tree based method bagging.  The random forest algorithm creates multiple trees from bootstrapped samples, averages those results, and uses a random subset of predictors for each bootstrap sample/tree fit.  Random forests can be used for classification and regression problems.

```{r, echo = TRUE}
rfFit <- train(shares ~., data = onlineNewsTrain, 
               method = "rf", 
               trainControl = trainControl(method = "cv",
                                           number = 5),
               tuneGrid = data.frame(mtry = 1:8))
rfFit
```

## Comparison

```{r, echo = TRUE}
# Compare linear regression model fit1 on the test set.
pred <- predict(fit1, newdata = onlineNewsTest)
postResample(pred, obs = onlineNewsTest$shares)

# Compare random forest model rfFit on the test set.
predForest <- predict(rfFit, newdata = onlineNewsTest)
postResample(predForest, onlineNewsTest$shares)
```
### Linear Regression Model

```{r}
# Set seed and split the data into 70% training and 30% test
set.seed(100)
index <- createDataPartition(onlineNewsLifestyle$shares, p = 0.7, list = FALSE)
train <- onlineNewsLifestyle[index,]
test <- onlineNewsLifestyle[-index,]

# Backward selection better than best subset for this dataset. 5 variables were chosen
model <- as.formula("shares ~ kw_avg_avg + kw_min_avg + kw_max_avg + num_hrefs + n_tokens_content")

# Model fit on training data
fit <- train(model, data = train,
              method = "lm",
              preProcess = c("center", "scale"),
              trControl = trainControl(method = "cv", number = 10))
```

### Boosted Tree Model 

```{r}
# Boosted tree. After examining ranges of values for number of trees and interaction depth, 25 trees with an interaction depth of 1 had the lowest test RMSE.
boostedTree <- train(shares ~. -data_channel_is_lifestyle, data = train,
             method = "gbm",
             preProcess = c("center", "scale"),
             trControl = trainControl(method = "repeatedcv", number = 5,
                                      repeats = 3),
             tuneGrid = expand.grid(n.trees = 25, interaction.depth = 1, shrinkage = 0.1, n.minobsinnode = 10))
```

## Comparisons

```{r}
# Test error computed for backwards selection linear model
pred <- predict(fit, newdata = test)
p <- postResample(pred, obs = test$shares)

# Test error computed for boosted tree model
pred <- predict(boostedTree, newdata = test)
p2 <- postResample(pred, obs = test$shares)

# Produce a table for all comparisons
comparisons <- t(rbind(p[1],p2[1]))
colnames(comparisons) <- c("Linear Model Backwards Selection", "Boosted Tree")
comparisons
```


## Automation



